---
title: "Recovering a True Model"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
p1Data <- read.csv("p1Data-1.csv", header=TRUE)
p2Data <- read.csv("p2Data-1.csv", header=TRUE)
p3Data <- read.csv("p3Data-1.csv", header=TRUE)
```

## Part 1 - Recovering a True Model (Beginner Level)

We believe the true model to be...

$$
  Y_i = \underbrace{\beta_0 + \beta_1 X_{4i} + \beta_2 X_{4i} X}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as 13.11. 

```{r}
pairs(p1Data)
pairs(p1Data, panel = panel.smooth, col = as.factor(p1Data$X2))

mylm1 <- lm(Y~X4, data = p1Data)
summary(mylm1)

plot(mylm1$residuals ~ ., data = p1Data, col = as.factor(p1Data$X2))
# this period is like SQL's *. It selects all columns one at a time in this case)

mylm2 <- lm(Y~X4 + X2 + X4:X2, data = p1Data)
summary(mylm2)




mylm1 <- lm(Y~X2, data = p1Data)
summary(mylm1)

plot(mylm1$residuals ~ ., data = p1Data, col = as.factor(p1Data$X2))
# this period is like SQL's *. It selects all columns one at a time in this case)

mylm2 <- lm(Y~X4 + X2, data = p1Data)
summary(mylm2)


# 
pairs(cbind(Res = mylm2$residuals, fit = mylm2$fitted.values, p1Data), panel=panel.smooth, col = as.factor(p1Data$X2))

mylm3 <- lm(Y~X4 + X2 + X3, data = p1Data)
summary(mylm3)
# remember, is a variable is not usless is it's not significant. Usefulness at this point in the model at predicting Y. 


mylm4 <- lm(Y~X4 + X2 + X4:X2, data = p1Data)
summary(mylm4)
# remember, interactions are what let things tilt/ 


mylm5 <- lm(Y~X4 + X4:X2, data = p1Data)
summary(mylm5)


```

Y:X2 & Y:X4 have decent patterns. 



<br>

## Part 2 - Recovering a True Model (Intermediate Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```



## Part 3 - Recovering a True Model (Advanced Level)

We believe the true model to be...

$$
  Y_i = \underbrace{...}_\text{The True Model} + \epsilon_i
$$

And estimate that model by...

$$
  \hat{Y}_i = ...
$$

with our estimate of $\sigma$ as ...

```{r}

```





Steps for Regression Battleship:
1. start with pairs() plot
2. look for patterns 
3. then go for the plot(mylm1$residuals ~ ., p1data) and look for thing
4. edit the lm based on your previous findings